Illuminating a Dark Lens : A Type Ia Supernova Magnified by Galaxy Cluster Abell 2744

### 2015-01-16 snTomas.tex
discovery and follow-up

\ifms
    \author{Steven A.~Rodney\altaffilmark{1}}
    \author{\etal\altaffilmark{2}}

    \altaffiltext{1}{\JHU}
    \altaffiltext{1}{\elsewhere}
\else 
    \author{
      Steven~A.~Rodney\altaffilmark{1},
      \etal\altaffilmark{2}
    }
    \altaffiltext{1}{\JHU}
    \altaffiltext{2}{\elsewhere}
\fi

@@

Lensed SNe have been discovered
\textcolor{red}{\citep{Goobar:2009,Riehm:2011,Patel:2014,Nordin:2014,}}

\section{Discovery and Follow-up}
\label{sec:DiscoveryAndFollowUp}

>>>>

\author{
 Steven~A.~Rodney\altaffilmark{1},
  \etal
}
\altaffiltext{1}{\JHU}
\altaffiltext{2}{Hubble fellow}

@@

Lensed SNe have been discovered
\citep{Goobar:2009,Riehm:2011,Patel:2014,Nordin:2014}

In Section~\ref{sec:DiscoveryAndFollowup} we present the discovery and
follow-up observations of SN \tomas at $z=1.31$, discovered behind the
galaxy cluster Abell 2744.  Sections~\ref{sec:Spectrum} and \ref{sec:LightCurve}
describe the spectroscopy and photometry of this SN, leading to a
classification of the object as a normal Type Ia SN.  In
Section~\ref{sec:LensingMagnification} we make a direct measurement of
the magnification of this source due to gravitational lensing, and
compare to predicitions from lens models.  Finally,
Section~\ref{sec:Discussion} discusses the tension between our
magnification measurement and the lens models, with implications for
the systematic error budget in magnification estimates for other
high-$z$ lensed sources.

\section{Discovery and Follow-up}
\label{sec:DiscoveryAndFollowup}

SN \tomas\ was discovered in \HST\ observations with the Advanced
Camera for Surveys (ACS) in the F606W and F814W bands (V and i),
collected on UT 2014 May 15 as part of the Hubble Frontier Fields
(HFF) survey (PI:J.Lotz,
HST-PID:13495).\footnote{\url{http://www.stsci.edu/hst/campaigns/frontier-fields}}
The HFF program is a 3-year director's discretionary initiative that
is collecting 140 orbits of HST imaging (roughly 340 ksec) on six
massive galaxy clusters, plus 6 accompanying ``parallel fields.''
Each field is observed in 3 optical bands (ACS F435W, F606W and F814W)
and 4 infrared (IR) bands (WFC3-IR F105W, F125W, F140W, and F160W),
although the optical and IR imaging campaigns are separated by $\sim$6
months. Abell 2744 was the first cluster observed, with IR imaging
spanning October-November 2013, and the optical imaging from May-July
2014.

The SN detection was made in difference images using template imaging
of Abell 2744 from HST+ACS observations taken in 2009 (PI:Dupke,
HST-PID:11689).  Upon discovery, HST target-of-opportunity
observations were triggered from the FrontierSN program (PI:Rodney,
HST-PID:13386), which aims to discover and follow transient sources in
the HFF cluster and parallel fields. The FrontierSN observations
provided WFC3-IR imaging as well as spectroscopy using the ACS G800L
grism, supplementing the rapid-cadence optical imaging from HST+ACS
already being provided by the HFF program. Difference images for the
IR follow-up data were generated using templates constructed from the
HFF WFC3-IR imaging campaign, which concluded in November, 2013.

All of the imaging data were processed using the {\tt sndrizpipe}
pipeline,\footnote{\url{https://github.com/srodney/sndrizpipe}
v1.2 DOI:10.5281/zenodo.10731} a custom data reduction package in
Python that is employs the {\tt DrizzlePac} tools from the Space
Telescope Science Institute (STScI) \citep{Fruchter:2010}.  Photometry
was collected using the {\tt PyPhot} software
package,\footnote{\url{https://github.com/djones1040/PyPhot}} a pure-Python
implementation of the photometry algorithms from the IDL AstroLib
package \citep{Landsman:1993}, which in turn are based on the DAOPHOT
program \citep{Stetson:1987}.  For the IR bands we used point spread
function (PSF) fitting on the difference images, and for the ACS
images we used 0\farcs3-aperture photometry.
